### 2020-02-29 doc/icfp20/intro.tex
add diagrams

Edgar Var\`{e}se describes music as ``organized sound'', and
throughtout history cultures have created and applied systems of rules
and guidelines to govern the music they create, most notably the
Common Practice Period of Western music spanning the 17th to early
20th centuries. These rule systems are seldom absolute, and indeed
deliberate breaking of the rules is often part of the aesthetic, but they
roughly constrain the music they apply to and
give it a common form and sound.

@@

We would like to explore what can be done in the context of music by
using a programming language that offers full dependent types. We use
the language Agda~\citep{norell-phd} since it is fairly mature and
aims to be a functional programming language as well as a proof
assistant. It also features a Haskell FFI so we can take advantage of
existing Haskell libraries, in partiular for MIDI.

Full dependent types allow expression of full predicate logic, and it is
tempting to take a standard textbook on music theory such as
\citet{piston-harmony} or \citet{aldwell2018harmony} and attempt to
formalize it in type theory. However this is difficult since most
``rules'' in music theory are not absolute, but rather suggestions of
varying degrees (themselves vaguely specified) of importance. Since
ultimately the resulting created music is more important than the
theory, an alternative is to attempt to indirectly deduce the rules
from music samples, using for example machine
learning~\citep{huang-cp}. However.... TODO
TODO: Perhaps create rules and weight them, then use ML.

@@

We make use of and extend~\cite{MusicTools}, a library of small tools
for analysis and synthesis of music written in Agda, with the goal
of eventually being a dependently-typed alternative to Euterpea. 
Previous work by ~\cite{cong-cp} using Music Tools expressed
first-species counterpoint with the Agda type system, so that an error
in counterpoint resulted in an Agda type error. Although this takes
advantage of Agda's native type checking, it has several
downsides. One is that the Agda errors may be difficult to interpret
as the corresponding musical error. Another is TODO: flexibility and
composability.

This is an experience report, and we highlight both the advantages and
disadvantages of using Agda

>>>>

Edgar Var\`{e}se describes music as ``organized sound'', and
throughout history cultures have created and applied systems of rules
and guidelines to govern the music they create, most notably the
Common Practice Period of Western music spanning the 17th to early
20th centuries. These rule systems are seldom absolute, and indeed
deliberate breaking of the rules is often part of the aesthetic, but they
roughly constrain the music they apply to and
give it a common form and sound.

@@

In this paper we explore what can be done in the context of music by
using a programming language that offers full dependent types. We use
the language Agda~\citep{norell-phd} since it is fairly mature and
aims to be a functional programming language as well as a proof
assistant. It also features a Haskell FFI so we can take advantage of
existing Haskell libraries, in partiular for MIDI.

Full dependent types allow expression of predicate logic, and it is
tempting to take a standard textbook on music theory such as
\citet{piston-harmony} or \citet{aldwell2018harmony} and attempt to
formalize it in type theory. However this is difficult since most
``rules'' in music theory are not absolute, but rather suggestions of
varying degrees (themselves vaguely specified) of importance. Since
ultimately the resulting created music is more important than the
theory, an alternative is to attempt to indirectly deduce the rules
from music samples, using for example machine
learning~\citep{huang-cp}. However.... TODO
TODO: Perhaps create rules and weight them, then use ML.

@@

We make use of and extend~\citet{MusicTools}, a library of small tools
for analysis and synthesis of music written in Agda, with the goal
of eventually being a dependently-typed alternative to Euterpea. 
Previous work by ~\citet{cong-cp} using Music Tools expressed
first-species counterpoint with the Agda type system, so that an error
in counterpoint resulted in an Agda type error. Although this takes
advantage of Agda's native type checking, it has several
downsides. One is that the Agda errors may be difficult to interpret
as the corresponding musical error. Another is TODO: flexibility and
composability.

This is an experience report, and we highlight both the advantages and
disadvantages of using Agda. TODO: MORE

TODO: Talk about conclusion.
